---
title: "CitiBike NYC — Geospatial Trip Analysis"
subtitle: "Plataformas de Analítica de Negocios para Organizaciones (Gpo 101) — Tec de Monterrey CSF"
author: "Ludovic Delot Bravo"
date: "2024-08-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Geospatial exploration of a CitiBike (NYC bike-share) trip dataset: trip
duration cleanup, per-rideable-type averages, and two `sf`/`tigris`-based maps
of trip start locations over NYC census blocks.

Data source: CitiBike system trip data (public, published monthly by
Lyft/CitiBike at https://citibikenyc.com/system-data). The raw CSV is not
included in this repo — see the README for the download link. Update
`DATA_PATH` below before running.

```{r libraries}
pacman::p_load(dplyr, hms, ggplot2, tidyr, sf, rnaturalearth, rnaturalearthdata,
                tigris, RColorBrewer)
```

```{r load-data}
DATA_PATH <- "data/citibike/202403-citibike-tripdata.csv"
df <- read.csv(DATA_PATH)

missing_data <- colSums(is.na(df))

df$tripduration <- difftime(df$ended_at, df$started_at, units = "mins")
hist(as.numeric(df$tripduration))
```

```{r filter-durations}
# Trips under a minute are usually a false start/re-dock; trips over 45
# minutes are typically lost/stolen bikes rather than genuine rides.
df <- filter(df, tripduration <= 45, tripduration > 0)
hist(as.numeric(df$tripduration))
```

```{r inspect}
head(df, 5)
colnames(df)
```

```{r average-duration}
df$tripduration <- as.numeric(df$tripduration)
promedio_duracion <- mean(df$tripduration, na.rm = TRUE)
promedio_duracion
```

```{r station-counts}
num_estaciones_inicio <- length(unique(df$start_station_name))
num_estaciones_fin    <- length(unique(df$end_station_name))
total_estaciones <- length(unique(c(df$start_station_name, df$end_station_name)))
total_estaciones
```

```{r export}
write.csv(df, file = "df.csv", row.names = FALSE)
```

```{r duration-by-bike-type}
tapply(df$tripduration, df$rideable_type, mean)
```

## Map 1: trip starts over a US basemap

```{r world-basemap}
world <- ne_countries(scale = "medium", returnclass = "sf")
usa <- subset(world, world$sov_a3 == "USA")

tabla1 <- df %>%
  group_by(start_station_id) %>%
  summarise(conteo = n(),
            longitud = mean(start_lng),
            latitud = mean(start_lat))

ggplot(data = usa) +
  geom_sf() +
  geom_point(data = tabla1, aes(x = longitud, y = latitud), size = 2,
             shape = 23, fill = "red") +
  coord_sf(xlim = c(-73.84, -74.05), ylim = c(40.63, 40.892), expand = FALSE)
```

## Map 2: trip density over NYC census blocks

```{r nyc-blocks}
nyc_blocks <- blocks(state = "NY", county = c("New York", "Kings", "Bronx", "Queens"),
                      year = 2020, class = "sf")
```

```{r nyc-block-map}
ggplot(data = nyc_blocks) +
  geom_sf(fill = "antiquewhite") +
  geom_point(data = tabla1, aes(x = longitud, y = latitud, size = conteo, fill = conteo),
             pch = 21, alpha = 0.7) +
  scale_fill_gradientn(colors = c("red", "yellow", "green")) +
  coord_sf(xlim = c(-73.84, -74.05), ylim = c(40.63, 40.892), expand = FALSE) +
  guides(size = "none") +
  theme(panel.background = element_rect(fill = "lightblue"))
```
