---
title: "Ecommerce Customers — Linear Regression"
subtitle: "Plataformas de Analítica de Negocios para Organizaciones (Gpo 101) — Tec de Monterrey CSF"
author: "Ludovic Delot Bravo"
date: "2024-08-21"
output: html_document
---

## Overview

Small regression exercise on a synthetic "Ecommerce Customers" dataset:
correlation/pairs exploration, four simple linear plots against
`yearly_amount_spent`, a multiple linear regression, and stepwise (AIC)
variable selection.

Cleanup notes:

- Fixed `aes(apha = 0.01)` -> `aes(alpha = 0.01)` in the `ggpairs()` call (typo
  silently ignored the intended transparency setting).
- Fixed `dev.off` (missing parentheses, so it never actually closed the
  graphics device) -> `dev.off()`.

Data source: this is the standard "Ecommerce Customers" teaching dataset used
in several R regression tutorials (columns: email, address, avatar,
avg_session_length, time_on_app, time_on_website, length_of_membership,
yearly_amount_spent). Not included here — see README.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries-and-load}
pacman::p_load(GGally, janitor, readr, dplyr, MASS)

DATA_PATH <- "data/ecommerce_customers.csv"
usuarios <- read_csv(DATA_PATH)

head(usuarios, 5)
usuarios <- clean_names(usuarios)
summary(usuarios)

# Drop identifying / non-predictive columns
usuarios <- usuarios %>%
  dplyr::select(-email, -address, -avatar)

hist(usuarios$yearly_amount_spent)
# Modeling assumptions to check before fitting: correlation, normality, linearity
```

```{r scatterplots}
plot(yearly_amount_spent ~ avg_session_length, data = usuarios)
plot(yearly_amount_spent ~ time_on_app, data = usuarios)
plot(yearly_amount_spent ~ time_on_website, data = usuarios)
plot(yearly_amount_spent ~ length_of_membership, data = usuarios)
```

```{r scatterplots-with-trend-lines}
plot(yearly_amount_spent ~ avg_session_length, data = usuarios,
     main = "Gasto Anual vs. Tiempo en Sesión",
     xlab = "Tiempo en Sesión", ylab = "Gasto Anual")
abline(lm(yearly_amount_spent ~ avg_session_length, data = usuarios), col = "red")

plot(yearly_amount_spent ~ time_on_app, data = usuarios,
     main = "Gasto Anual vs. Tiempo en la App",
     xlab = "Tiempo en la App", ylab = "Gasto Anual")
abline(lm(yearly_amount_spent ~ time_on_app, data = usuarios), col = "red")

plot(yearly_amount_spent ~ time_on_website, data = usuarios,
     main = "Gasto Anual vs. Tiempo en el Sitio Web",
     xlab = "Tiempo en el Sitio Web", ylab = "Gasto Anual")
abline(lm(yearly_amount_spent ~ time_on_website, data = usuarios), col = "red")

plot(yearly_amount_spent ~ length_of_membership, data = usuarios,
     main = "Gasto Anual vs. Duración de Membresía",
     xlab = "Duración de Membresía", ylab = "Gasto Anual")
abline(lm(yearly_amount_spent ~ length_of_membership, data = usuarios), col = "red")

dev.off()
cor(usuarios)
ggpairs(usuarios, mapping = aes(alpha = 0.01))
```

```{r regression-and-stepwise-selection}
spent <- lm(yearly_amount_spent ~ length_of_membership + avg_session_length +
              time_on_app + time_on_website, data = usuarios)
summary(spent)

# Stepwise (AIC-based) variable selection
modelo <- stepAIC(spent, direction = "both", trace = FALSE)
summary(modelo)

newdata <- data.frame(
  length_of_membership = mean(usuarios$length_of_membership),
  avg_session_length   = mean(usuarios$avg_session_length),
  time_on_app          = mean(usuarios$time_on_app),
  time_on_website       = mean(usuarios$time_on_website))

y_predict <- predict(modelo, newdata)
y_predict
```
